---
title: "3. EDA"
author: "Giorgio Rivoli"
date: "2024-08-09"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

Let’s now proceed with the exploratory data analysis.
We will start with a descriptive analysis of the variables present in the dataset, and then perform some statistical tests.

```{r}
# Load necessary libraries
library(readxl)  # For reading Excel files
library(openxlsx)     # For writing Excel files
library(dplyr)  # For data manipulation (part of the tidyverse)
library(tidyr)  # For tidying data (part of the tidyverse)
library(tidyverse)  # A collection of R packages for data science (includes dplyr, ggplot2, tidyr, etc.)
library(ggplot2)  # For data visualization
library(corrplot)  # For correlation matrix visualization
library(purrr)  # For functional programming (part of the tidyverse)
library(lubridate)  # For working with dates and times
library(stats)  # Base R statistical functions
library(car)  # For Levene’s test (used to assess the equality of variances)
library(effectsize)  # For calculating effect sizes
library(effsize)  # Another package for effect size calculations
library(gridExtra)  # For arranging multiple plots in a grid
```

```{r}
# Read in the Excel file containing the movie dataset
movie_data <- read_excel("movie_data.xlsx")

# Check for missing values in the entire dataset
sum(is.na(movie_data))
# Check for missing values in the Director_gender column
sum(is.na(movie_data$Director_gender))
# Check for missing values in the Screenplay_gender column
sum(is.na(movie_data$Screenplay_gender))
```

Upon rechecking for NA values in the dataset, we found 3481 missing values.
Previously, these NAs were not visible because the code left a blank space in case of missing gender, instead of inserting an NA.
This issue concerns the absence of the director’s gender in 34 cases and the screenwriter’s gender in 3447 cases.
What we are going to do is removing the observations where there is an NA for the `Director_gender` variable.
We don't do the same for the screenwriters because, at the end, we are going to remove the variables related to this role.

```{r}
movie_data <- movie_data |> 
  filter(!is.na(movie_data$Director_gender))
```

# Descriptive Analysis

Regarding the descriptive analysis, we will proceed by examining the variables, starting from the first one and progressively moving to the right.

## Title

### Mean

For these two variables, we will exclusively analyze the average number of words in the original and translated titles.

```{r}
# Calculate and print the mean number of words in the original title
mean_nmb_wrd_orgnl = mean(movie_data$nmb_wrd_orgnl, na.rm = TRUE)
cat("The average number of words in the original title is:", round(mean_nmb_wrd_orgnl, 2), "\n")

# Calculate and print the mean number of words in the translated title
mean_nmb_wrd_title = mean(movie_data$nmb_wrd_title, na.rm = TRUE)
cat("The average number of words in the title is:", round(mean_nmb_wrd_title, 2), "\n")

# Remove the mean variables from the environment to free up space
rm(mean_nmb_wrd_orgnl)
rm(mean_nmb_wrd_title)
```

## Title Sentiment

### Mean

Regarding the sentiment analysis of the title, we will start by examining the average AFINN score.

```{r}
# Calculate and print the mean sentiment analysis score for the title (using the AFINN method)
mean_afinn_title = mean(movie_data$afinn_title, na.rm = TRUE)
cat("The average sentiment analysis for the title is:", mean_afinn_title)

# Remove the mean_afinn_title variable after use
rm(mean_afinn_title)
```

As we might expect, the average Afinn scores are close to 0 (slightly negative).
This indicates that, on average, the headlines contain neutral words.

### vs Revenue/Budget Ratio

Let’s delve deeper into the sentiment analysis of the title.
We’ll start by graphically visualizing the relationship between the title’s sentiment and the ratio of a film’s revenue to its budget.

```{r}
# Plot the relationship between title sentiment and the revenue/budget ratio
ggplot(movie_data, aes(x = as.numeric(afinn_title), y = rev_bud_ratio)) +
  geom_point(color = "purple") +
  labs(x = "Title Sentiment", y = "Revenue-Budget Ratio", 
       title = "Title Sentiment vs Revenue/Budget Ratio") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center and style the plot title
    axis.title = element_text(size = 10)  # Adjust axis title font size
  )
```

We observe that films with the highest ratio tend to have a neutral sentiment score, or at least not excessively negative or positive.

### vs Year

Let’s now examine the evolution of the sentiment of the titles over time.
This will allow us to understand if there have been significant changes in the titles themselves, particularly if more impactful titles have been chosen from a communicative point of view.

```{r}
# Plot the trend of title sentiment over time (by year)
ggplot(movie_data, aes(x = as.numeric(year), y = afinn_title)) +
  geom_point() +
  labs(x = "Year", y = "Title Sentiment", title = "Title Sentiment trend by Year") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.title = element_text(size = 10)  # Adjust axis title font size
  )
```

Although not markedly, it is observed that the band of points tends to be a little bit wider over time.
This suggests that, over the years, there has been an increasing number of titles with strong communicative effectiveness.
It is likely that in the past there was greater control, perhaps even censorship, over the titles, which has gradually eased.
Additionally, people have generally become more accustomed to strong sentiments associated with audiovisual works.

### vs Certification

Let’s now examine the relationship between the sentiment of the title and the rating of a film.

```{r}
# Plot the relationship between title sentiment and movie certification
ggplot(movie_data, aes(x = factor(certification), y = afinn_title)) +
  geom_jitter(aes(color = afinn_title), alpha = 0.6, width = 0.2) +  # Add jittered points colored by sentiment
  geom_boxplot(alpha = 0.3, outlier.shape = NA) +  # Add transparent boxplots without outliers
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white") +  # Highlight the mean for each certification
  scale_x_discrete(labels = c("G", "PG", "PG-13", "R", "NC-17", "NR")) +  # Rename certification labels
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +  # Color scale for sentiment
  labs(x = "Certification", y = "Title Sentiment", 
       title = "Title Sentiment by Certification",
       color = "Sentiment") +
  theme_minimal() +
  theme(legend.position = "right") +  # Place legend on the right
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center the plot title
    axis.title = element_text(size = 10)  # Adjust axis title font size
  )
```

It is observed that films with the lowest rating, that is, those suitable for all ages, generally have a title sentiment close to neutral.
On the contrary, as the rating becomes more restrictive, the variability of the title sentiments increases.
This suggests that films intended for a more mature audience can afford to use language that evokes stronger emotions.

## Budget

### Mean

Also in this case, let’s start by analyzing the average of the budget variable.

```{r}
# Calculate the mean budget while ignoring missing values (NA)
mean_budget = mean(movie_data$budget, na.rm = TRUE)
cat("The average budget is:", mean_budget, "\n")

# Remove the mean_budget variable after use
rm(mean_budget)
```

The average budget is around 24 million dollars.

### Distribution

Let’s now examine the distribution of this variable.

```{r}
# Plot the distribution of movie budgets
ggplot(movie_data, aes(x = budget)) +
  geom_area(stat = "bin", bins = 50, fill = "darkblue", alpha = 1) +  # Use an area plot to show the budget distribution
  scale_x_continuous(labels = scales::comma_format(scale = 1e-6, suffix = "M"),
                     breaks = seq(0, 400e6, by = 100e6)) +  # Format x-axis in millions with "M" as the suffix
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = seq(0, 3500, by = 1000)) +  # Format y-axis with commas
  labs(title = "Budget Distribution",
       x = "Budget ($)", 
       y = "Number of Movies") +
  theme_minimal() +  # Use a minimal theme for a clean look
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center the title and style it
    axis.title = element_text(size = 10),  # Adjust axis title font size
    axis.text = element_text(size = 8)  # Adjust axis text font size
  ) +
  coord_cartesian(xlim = c(0, 400e6), ylim = c(0, 3500))  # Set limits for x and y axes to focus the plot
```

As expected, most films have a budget that is below or close to the average.
As the budget increases, the number of films produced decreases.

### vs Revenue

Now let’s analyze the relationship between a film’s budget and its revenue.

```{r}
# Plot the relationship between budget and revenue
ggplot(movie_data, aes(x = budget, y = revenue)) +
  geom_point(alpha = 0.4, color = "darkred", size = 1.5) +  # Add scatterplot points
  geom_smooth(method = "lm", color = "black", se = FALSE, size = 1) +  # Add a linear regression line without confidence intervals
  scale_x_continuous(labels = scales::comma_format(scale = 1e-6, suffix = "M"),
                     breaks = seq(0, 500e6, by = 100e6)) +  # Format x-axis in millions with "M" suffix
  scale_y_continuous(labels = scales::comma_format(scale = 1e-6, suffix = "M"),
                     breaks = seq(0, 3e9, by = 500e6)) +  # Format y-axis in billions with "M" suffix
  labs(title = "Budget vs Revenue",
       x = "Budget ($)",
       y = "Revenue ($)") +
  theme_minimal() +  # Use a minimal theme for a clean look
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center the title and style it
    plot.subtitle = element_text(hjust = 0.5, size = 12, color = "darkgrey"),  # Style the subtitle
    axis.title = element_text(size = 10),  # Adjust axis title font size
    axis.text = element_text(size = 8),  # Adjust axis text font size
    panel.grid.minor = element_blank()  # Remove minor grid lines for clarity
  ) +
  coord_cartesian(ylim = c(0, 3e9), xlim = c(0, 450e6))  # Set limits for x and y axes to focus the plot
```

In the graph, the red line represents the trend or regression line, which highlights the general relationship between the budget and the revenue.
The positive slope of the line indicates that, generally, as the budget increases, the revenue also increases.
This line provides a visual estimate of the correlation between the two variables, allowing for a quick observation of the overall trend of the data, represented by the blue points in the graph.

## Revenue

### Mean

Below is the code to calculate the average revenue.

```{r}
# Calculate the mean revenue while ignoring missing values (NA)
mean_revenue = mean(movie_data$revenue, na.rm = TRUE)
cat("The average revenue is:", mean_revenue, "\n")

# Remove the mean_revenue variable after use
rm(mean_revenue)
```

The average revenue of a film is approximately 67 million dollars.
As expected, this figure is significantly higher than the average budget.

### Distribution

Now let’s analyze the distribution of this variable:

```{r}
# Plot the distribution of movie revenues
ggplot(movie_data, aes(x = revenue)) +
  geom_area(stat = "bin", bins = 50, fill = "darkred", alpha = 1) +  # Use an area plot to show the revenue distribution
  scale_x_continuous(labels = scales::comma_format(scale = 1e-6, suffix = "M"),
                     breaks = seq(0, 1200e6, by = 100e6)) +  # Format x-axis in millions with "M" as the suffix
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = seq(0, 6500, by = 1000)) +  # Format y-axis with commas
  labs(title = "Revenue Distribution",
       x = "Revenue ($)", 
       y = "Number of Movies") +
  theme_minimal() +  # Use a minimal theme for a clean look
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center the title and style it
    axis.title = element_text(size = 10),  # Adjust axis title font size
    axis.text = element_text(size = 8)  # Adjust axis text font size
  ) +
  coord_cartesian(xlim = c(0, 1200e6), ylim = c(0, 6500))  # Set limits for x and y axes to focus the plot
```

Similarly to what was observed for the budget variable, we also notice that most films have a revenue that is below or close to the average.
Moving to the right in the graph, that is, examining the number of films with very high revenues, we observe a rapid decrease in the number of films.

## Revenue/Budget Ratio

### Mean

```{r}
# Calculate the mean revenue-to-budget ratio while ignoring missing values (NA)
mean_rev_bud_ratio = mean(movie_data$rev_bud_ratio, na.rm = TRUE)
cat("The average revenue/budget ratio is:", round(mean_rev_bud_ratio, 2))

# Remove the mean_rev_bud_ratio variable after use
rm(mean_rev_bud_ratio)
```

We observe that the average value of the ratio between revenue and budget is 5.76.
This means that, on average, a film earns around 5 times its budget.
It is evident that this is a very satisfactory result.

### vs Year

Now let’s analyze the trend of the ratio between revenue and budget over time.

```{r}
# Ensure the 'year' column is treated as numeric
movie_data <- movie_data |> 
  mutate(year = as.numeric(year))

# Plot the revenue-to-budget ratio over time (by year)
ggplot(movie_data, aes(x = year, y = rev_bud_ratio)) +
  geom_point(color = "purple") +  # Add scatterplot points
  scale_x_continuous(breaks = seq(min(movie_data$year), 
                                  max(movie_data$year), 
                                  by = 10)) +  # Set x-axis breaks every 10 years
  labs(title = "Revenue/Budget Ratio Over Time",
       x = "Year",
       y = "Revenue / Budget") +
  theme_minimal() +  # Use a minimal theme for a clean look
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center the title and style it
    axis.title = element_text(size = 10),  # Adjust axis title font size
    axis.text = element_text(size = 8)  # Adjust axis text font size
  )
```

It can be observed that, generally, over the years, this ratio has increased, with some films recording extremely high values.
The reasons for this phenomenon can be manifold.
Firstly, over time, cinema has become an increasingly well-known and accessible service.
Another explanation could lie in the emergence of new mass media, such as television first and then the internet, which have made advertising and word-of-mouth, and therefore virality, much more effective, thus increasing the number of viewers for a particular film.

## Date

### Film count per month

Let’s start by counting the number of films released each month.

```{r}
# Number of movies per month
month_names <- c("January", "February", "March", "April", "May", "June", 
                 "July", "August", "September", "October", "November", "December")

# Count the number of movies released each month
films_per_month <- movie_data |> 
  count(month) |> 
  arrange(month) |> 
  mutate(month_name = factor(month_names, levels = month_names))

# Plot the number of movies released per month
ggplot(films_per_month, aes(x = n, y = month_name)) +
  geom_col(fill = "skyblue", color = "black") +  # Create a bar plot with skyblue bars
  theme_minimal() +  # Use a minimal theme for a clean look
  labs(title = "Number of Movies per Month",
       x = "Number of Movies",
       y = "Month") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center and style the title
        axis.text.y = element_text(hjust = 1, face = "italic")) +  # Italicize the y-axis text
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, max(films_per_month$n) * 1.1),  # Set x-axis limits to be slightly wider than max count
                     breaks = seq(0, 1200, by = 200))  # Set x-axis breaks every 200 units

# Remove unnecessary variables from the environment
rm(films_per_month)
rm(month_names)
```

We observe that the months with the highest number of film releases are the final months of the year.
The reasons can be various.
Firstly, during these months it starts to get colder, and going to the cinema can be a good alternative to other outdoor activities.
Additionally, during this period there are many holidays, such as Halloween and Christmas, which can lead to an increase in cinema attendance.

### Month vs Revenue

Now let’s analyze the relationship between the release month of a film and its earnings.

```{r}
month_names <- c("January", "February", "March", "April", "May", "June", 
                 "July", "August", "September", "October", "November", "December")

# Calculate the average revenue for each month and sort in descending order
monthly_revenue <- movie_data |> 
  group_by(month) %>%
  summarise(avg_revenue = mean(revenue, na.rm = TRUE)) %>%
  arrange(desc(avg_revenue))

# Convert the month column to a factor ordered by the month names
monthly_revenue$month <- factor(month_names[as.numeric(monthly_revenue$month)], 
                                levels = month_names[as.numeric(monthly_revenue$month)])

# Plot the average revenue by month
ggplot(monthly_revenue, aes(x = avg_revenue, y = month)) +
  geom_bar(stat = "identity", fill = "darkred") +  # Create a bar plot with darkred bars
  labs(title = "Average Revenue by Month",
       x = "Average Revenue",
       y = "Month") +
  theme_minimal() +  # Use a minimal theme for a clean look
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Center and style the title
        axis.text.y = element_text(hjust = 1, face = "italic")) +  # Italicize the y-axis text
  scale_x_continuous(labels = scales::label_number(suffix = "M", scale = 1e-6))  # Format x-axis labels in millions with "M" suffix

# Remove unnecessary variables from the environment
rm(monthly_revenue)
rm(month_names)
```

We observe that the most profitable months are June, July, and May, which are the summer months.
This dataset is mainly composed of American films, and in the USA, the summer period is one of the favorite times to go to the cinema.
This may also be due to the fact that, during the summer, people seek relief in air-conditioned cinemas, in a logic similar to that of winter releases.
In fact, right after these three months, we notice that the most profitable months are December and November, when it is colder and there are holidays like Christmas.

## IP

### Mean

Let’s start by analyzing the proportion of films that are derived from other works, such as books, short stories, or video games.

```{r}
# Calculate the mean proportion of movies stemming from other works (IP) and display it
mean_ip = mean(movie_data$IP, na.rm = TRUE)
cat("The ", round(mean_ip*100, 2), "% of movies in the dataset stems from other works", sep = "")

# Remove the mean_ip variable after use
rm(mean_ip)
```

### vs Year

Now let’s examine the evolution of films based on intellectual property (IP) over the year

```{r}
# Calculate the proportion of IP films by year
proportion_by_year <- movie_data |> 
  group_by(year) |> 
  summarise(proportion = mean(IP, na.rm = TRUE))

# Plot the evolution of the proportion of IP films over time
ggplot(proportion_by_year, aes(x = year, y = proportion)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", 
       y = "Proportion of IP films",
       title = "Evolution of the Proportion of IP Films Over Time") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = seq(min(proportion_by_year$year), 
                                  max(proportion_by_year$year), 
                                  by = 10))  # Set x-axis breaks every 10 years

# Remove the proportion_by_year variable after use
rm(proportion_by_year)
```

In the early 20th century, the representation is quite peculiar, mainly because the dataset contains very few films from that period.
However, we observe that, as time goes by, the proportion of films derived from other works has progressively decreased.

### vs Revenue

Let’s examine whether this gradual reduction is due to the fact that, on average, films based on intellectual property (IP) earn less compared to non-IP films.

```{r}
# Calculate the average revenue for IP and non-IP films
revenue_by_ip <- movie_data |> 
  group_by(IP) |> 
  summarise(mean_revenue = mean(revenue, na.rm = TRUE)) |> 
  mutate(ip_status = ifelse(IP == 1, "IP", "NOT IP"))

# Plot the average revenue per IP and non-IP film
ggplot(revenue_by_ip, aes(x = ip_status, y = mean_revenue, fill = ip_status)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(x = "IP Status", 
       y = "Average Revenue", 
       title = "Average Revenue per IP and Non-IP Film") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
  scale_y_continuous(labels = scales::label_number(suffix = "M", scale = 1e-6)) +
  scale_fill_manual(values = c("NOT IP" = "#7fcdbb", "IP" = "#2c7fb8")) +  # Set custom colors for IP status
  theme(legend.position = "none")  # Remove the legend

# Remove the revenue_by_ip variable after use
rm(revenue_by_ip)
```

Surprisingly, we notice that IP films earn more on average compared to non-IP films.
Therefore, the decline over time might be attributed to other causes, such as copyright issues and the high costs of acquiring a work.

## Overview Sentiment

We always start from the average.

```{r}
# Calculate the mean overview sentiment score and display it
mean_overview = mean(movie_data$overview_sentiment, na.rm = TRUE)
cat("The average overview sentiment is:", mean_overview)

# Remove the mean_overview variable after use
rm(mean_overview)
```

Similarly to titles, movie plots also tend to have an average score close to neutral, that is, around 0, although in this case the value deviates slightly more.

### Distribution

Now let’s examine the distribution of this variable.

```{r}
# Plot the distribution of overview sentiment scores
ggplot(movie_data, aes(x = overview_sentiment)) +
  geom_histogram(binwidth = 1, fill = "#31a354", alpha = 1) +
  labs(title = "Overview Sentiment Distribution",
       x = "Overview Sentiment", 
       y = "Number of Movies") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

The distribution is very close to a normal distribution.
In fact, we notice that most films have a sentiment associated with the plot that is neutral or not particularly strong.
This is predictable, as films with too intense emotions can be overwhelming or even cloying, thus discouraging people from going to the cinema.

### vs Year

Now let’s examine if and how the sentiment in movie plots has changed over the years.

```{r}
# Plot overview sentiment scores by year
ggplot(movie_data, aes(x = as.numeric(year), y = overview_sentiment)) +
  geom_point() +
  labs(x = "Year", 
       y = "Overview Sentiment", 
       title = "Overview Sentiment by Year") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

Similarly to what was observed for the sentiment of titles, movie plots have also started to express increasingly intense emotions over time.
This could be due to the fact that the audience has become accustomed to softer stories, preferring instead more gripping plots.

### vs Revenue/Budget Ratio

Now let’s compare the sentiment of the overview with the revenue-to-budget ratio.

```{r}
# Plot the relationship between overview sentiment and the revenue-to-budget ratio
ggplot(movie_data, aes(x = as.numeric(overview_sentiment), y = rev_bud_ratio)) +
  geom_point(color = "purple") +
  labs(x = "Overview Sentiment", 
       y = "Revenue-Budget Ratio", 
       title = "Revenue/Budget Ratio vs Overview Sentiment") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

We observe that films with a higher revenue-to-budget ratio tend to have a neutral overview or values that are not too extreme.

## Original Language

### Mean

For this variable, we analyze only the proportion of films with English as the original language.
The American market has the strongest film industry and also provides the most data on produced films.
Consequently, most of the films in our dataset come from the USA, and we expect the proportion of films with English as the original language to be skewed in favor of the latter.

```{r}
# Calculate the percentage of movies with English as the original language and display it
mean_original_language = round(mean(movie_data$original_language), 2) * 100
cat("The ", mean_original_language, "% of movies has English as the original language", sep = "")

# Remove the mean_original_language variable after use
rm(mean_original_language)
```

Indeed, the percentage is very high, as expected.

## Spoken Languages

### Mean

Now let’s analyze, on average, how many languages are spoken in a film.

```{r}
# Calculate the mean number of spoken languages per movie and display it
mean_spoken = round(mean(movie_data$nmb_spoken_languages, na.rm = TRUE), 2)
cat("The average number of spoken languages is:", mean_spoken)

# Remove the mean_spoken variable after use
rm(mean_spoken)
```

In fact, on average, fewer than two languages are spoken in a film.

### Revenue

Let’s now examine the relationship between the number of languages spoken in a film and its box office revenue.

```{r}
# Plot the relationship between the number of spoken languages and revenue
ggplot(movie_data, aes(x = nmb_spoken_languages, y = revenue)) +
  geom_point(alpha = 0.5, color = "darkred") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-6, suffix = "M"),
                     breaks = seq(0, 3e9, by = 500e6)) +
  scale_x_continuous(breaks = seq(0, max(movie_data$nmb_spoken_languages, na.rm = TRUE), by = 1)) +
  labs(title = "Number of Languages vs Revenue",
       x = "Number of Spoken Languages",
       y = "Revenue ($)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

We observe that, with the exception of some outliers, films in which fewer languages are spoken tend to have higher box office revenues.
This could be because a greater number of spoken languages requires more cognitive effort from the viewer.

## Runtime

### Mean

For the runtime variable as well, let’s start by analyzing its average.

```{r}
# Calculate the mean runtime and display it
mean_runtime = mean(movie_data$runtime, na.rm = TRUE)
cat("The average runtime in minutes is:", round(mean_runtime, 0))

# Remove the mean_runtime variable after use
rm(mean_runtime)
```

We observe that, on average, the films produced have a runtime slightly under two hours.

### Distribution

Now let’s analyze the distribution of the “runtime” variable.

```{r}
# Plot the distribution of runtimes
ggplot(movie_data, aes(x = runtime)) +
  geom_area(stat = "bin", bins = 50, fill = "#feb24c", alpha = 1) +
  labs(title = "Runtime Distribution",
       x = "Minutes", 
       y = "Number of Movies") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  ) 
```

As we can observe, most films have a runtime of less than 130 minutes, while only a few significantly exceed two hours.
These longer films are less accessible to the audience, as they require a greater time commitment to be watched in theaters.
Additionally, long-duration films often require greater cognitive effort and attention, increasing the risk of becoming boring.

### vs Certification

Let’s graphically examine the relationship between a film’s runtime and its rating.

```{r}
# Plot runtime versus certification categories
ggplot(movie_data, aes(x = factor(certification), y = runtime)) +
  geom_boxplot() +
  labs(title = "Runtime vs Certification",
       x = "Certification", 
       y = "Minutes") +
  scale_x_discrete(labels = c("1" = "G", "2" = "PG", "3" = "PG-13", "4" = "R", "5" = "NC-17", "6" = "NR")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

The image shows the distribution of film runtimes for different certifications.
We observe a general trend of increasing median film duration with higher certification numbers, with categories PG-13 and NC-17 having higher medians.
All categories have outliers, i.e., films with exceptionally long durations, but these are more evident in the higher categories.
Category G, presumably corresponding to children’s or family films, shows the lowest median duration and the least variability, suggesting greater uniformity in the length of these films.
The intermediate categories show a gradual progression in duration, while higher categories seem to have a wider distribution, indicating greater flexibility in runtime for films with these certifications.
This analysis suggests a correlation between the type of certification and the film’s duration, likely reflecting audience expectations and genre conventions associated with each rating category.

### vs Revenue

Now let’s analyze the relationship between the duration of a movie and its earnings.

```{r}
# Plot the relationship between runtime and revenue
ggplot(movie_data, aes(x = runtime, y = revenue)) +
  geom_point(alpha = 0.5, color = "darkred") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-6, suffix = "M"),
                     breaks = seq(0, 3e9, by = 500e6)) +
  scale_x_continuous(breaks = seq(0, max(movie_data$runtime, na.rm = TRUE), by = 50)) +
  labs(title = "Runtime vs Revenue",
       x = "Minutes",
       y = "Revenue ($)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

As we can easily observe, and as one might expect, the highest earnings are obtained from movies with a duration between 100 and 200 minutes.
When the duration significantly exceeds three hours, the movie’s earnings decrease considerably.

### vs Year

Now let’s examine how the duration of movies has changed over time.

```{r}
# Plot runtime trend over the years
ggplot(movie_data, aes(x = as.numeric(year), y = runtime)) +
  geom_point() +
  labs(x = "Year", y = "Minutes", title = "Runtime Trend by Year") +
  scale_x_continuous(breaks = seq(min(movie_data$year), 
                                  max(movie_data$year), 
                                  by = 10)) +
  scale_y_continuous(breaks = seq(min(movie_data$runtime, na.rm = TRUE),
                                  max(movie_data$runtime, na.rm = TRUE),
                                  by = 50)) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

We observe that, over the years, although not markedly, the number of movies with a longer duration has slightly increased.
The reasons for this phenomenon can be manifold, ranging from artistic to technical motivations.
For example, in the past, movies were shot exclusively on film, which was very expensive, and this limited the duration of movies.
Today, however, most movies are made in digital format, where the only limit to the duration is the digital memory capacity.

## Tagline

### (Sentiment) Distribution

First, let’s analyze the distribution of sentiment related to the tagline.

```{r}
# Plot the distribution of tagline sentiment scores
ggplot(movie_data, aes(x = tagline_afinn)) +
  geom_histogram(binwidth = 1, fill = "dodgerblue3", alpha = 1) +
  scale_x_continuous(breaks = seq(min(movie_data$tagline_afinn, na.rm = TRUE),
                                  max(movie_data$tagline_afinn, na.rm = TRUE),
                                  by = 1)) +
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = seq(0, 6000, by = 500)) +
  labs(title = "Tagline Sentiment Distribution",
       x = "Tagline Sentiment", 
       y = "Number of Movies") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  ) 
```

### vs Revenue

Similarly to what was done for the sentiment related to the overview, we also examine the relationship between the sentiment of the tagline and the revenue.

```{r}
# Plot the relationship between tagline sentiment and revenue/budget ratio
ggplot(movie_data, aes(x = as.numeric(tagline_afinn), y = rev_bud_ratio)) +
  geom_point(color = "purple") +
  labs(x = "Tagline Sentiment", 
       y = "Revenue-Budget Ratio", 
       title = "Revenue/Budget Ratio vs Tagline Sentiment") +
  scale_x_continuous(breaks = seq(min(movie_data$tagline_afinn, na.rm = TRUE),
                                  max(movie_data$tagline_afinn, na.rm = TRUE),
                                  by = 1)) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  ) 
```

In this case as well, movies with a more neutral tagline generally present a higher level of revenue.

## Collection

### Mean

Let’s start by analyzing the proportion of movies in our dataset that are part of a collection.

```{r}
# Calculate the percentage of movies belonging to a collection and display it
mean_collection = mean(movie_data$belongs_to_collection, na.rm = TRUE)
cat("The ", round(mean_collection * 100, 2), "% of movies in the dataset belongs to a collection", sep = "")

# Remove the mean_collection variable after use
rm(mean_collection)
```

### vs Year

Next, we examine how this proportion has changed over the years.

```{r}
# Calculate the proportion of movies belonging to a collection per year
proportion_by_year <- movie_data |> 
  group_by(year) |> 
  summarise(proportion = mean(belongs_to_collection, na.rm = TRUE))

# Create a line chart showing the evolution of the proportion of movies belonging to a collection over time
ggplot(proportion_by_year, aes(x = year, y = proportion)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", 
       y = "Proportion of films belonging to a collection",
       title = "Evolution of the proportion of films belonging to a collection over time") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = seq(min(proportion_by_year$year), 
                                  max(proportion_by_year$year), 
                                  by = 10)) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )

# Remove the proportion_by_year variable after use
rm(proportion_by_year)
```

Except for some values in the first half of the 20th century, where the high proportion is mainly due to the low number of movies produced in those years, we observe a general upward trend.
This result is quite logical: over time, the number of successful movies increases, and therefore producers are more inclined to start or continue sagas, as they represent a relative guarantee of profit.

### vs Revenue

Let’s examine whether the increase over time in movies belonging to collections is supported by an increase in revenue.

```{r}
# Calculate the average revenue for movies that belong to a collection and those that don't
revenue_by_collection <- movie_data |> 
  group_by(belongs_to_collection) |> 
  summarise(mean_revenue = mean(revenue, na.rm = TRUE)) |> 
  mutate(collection_status = ifelse(belongs_to_collection == 1, 
                                    "Belongs to collection", 
                                    "Doesn't belong to collection"))

# Create a bar chart comparing the average revenue between collection and non-collection movies
ggplot(revenue_by_collection, aes(x = collection_status, y = mean_revenue, fill = collection_status)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(x = "Collection Status", 
       y = "Average Revenue", 
       title = "Average revenue per collection and non-collection movies") +
  theme_minimal() +
  scale_y_continuous(labels = scales::label_number(suffix = "M", scale = 1e-6)) +
  scale_fill_manual(values = c("Doesn't belong to collection" = "#c994c7", 
                               "Belongs to collection" = "#dd1c77")) +
  theme(legend.position = "none")

# Remove the revenue_by_collection variable after use
rm(revenue_by_collection)
```

As expected, the average income of movies that are part of a collection is significantly higher compared to those that are not.
In fact, movies in a saga, although not having a mathematical certainty, have a much higher probability of achieving guaranteed earnings, as they can rely on a loyal viewer base.

## Production Countries

### Mean

As previously observed for the variable *original language*, a predominance of films produced in the **United States** was expected.

```{r}
# Calculate the percentage of movies produced in the USA
mean_country = round(mean(movie_data$production_countries, na.rm = TRUE), 2) * 100
cat("The ", mean_country, "% of movies is produced in USA", sep = "")

# Remove the mean_country variable after use
rm(mean_country)
```

The results obtained fully confirm these expectations.

## Certification

### Frequency

Let’s start by analyzing the number of films associated with each rating.

```{r}
# Plot the frequency of movie ratings
movie_data |> 
  count(certification) |> 
  mutate(
    certification = factor(certification, levels = c("1", "2", "3", "4", "5", "6"),
                           labels = c("G", "PG", "PG-13", "R", "NC-17", "NR"))
  ) |> 
  ggplot(aes(x = certification, y = n)) +
  geom_col(fill = "skyblue") +
  labs(title = "Frequency of Movie Ratings",
       x = "Rating",
       y = "Number of Movies") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

It is observed that the rating with the highest number of films is the **R** rating, which stands for ***restricted***.

### vs Revenue/Budget Ratio

Let’s now examine whether R-rated films have the best revenue-to-budget ratio.
If this were the case, it would immediately explain why so many R-rated films are produced.

```{r}
# Calculate the average revenue/budget ratio by movie certification
certification_revenue <- movie_data |> 
  group_by(certification) |> 
  summarise(avg_rev_bud_ratio = mean(rev_bud_ratio, na.rm = TRUE)) |> 
  mutate(certification = factor(certification,
                                levels = c("1", "2", "3", "4", "5", "6"),
                                labels = c("G", "PG", "PG-13", "R", "NC-17", "NR")))

# Create a bar chart showing the average revenue/budget ratio by certification
ggplot(certification_revenue, aes(x = certification, y = avg_rev_bud_ratio)) +
  geom_bar(stat = "identity", fill = "purple") +
  scale_y_continuous(breaks = seq(min(movie_data$rev_bud_ratio),
                                  max(movie_data$rev_bud_ratio),
                                  by = 10)) +
  labs(title = "Average Revenue/Budget Ratio by Certification",
       x = "Certification",
       y = "Average Revenue/Budget Ratio") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )

# Remove the certification_revenue variable after use
rm(certification_revenue)
```

We observe that R-rated films have a lower revenue-to-budget ratio compared to NC-17 and G-rated films.
NC-17 films are often low-budget, like many horror films, which can rely on word-of-mouth and virality, as they are often associated with strong emotions like fear.
This value could be determined by a few films that, with a very low budget, have become real success stories, earning a lot.
On the other hand, the second genre with the best revenue-to-budget ratio is easily explained: G-rated films.
These films can be watched by people of all ages, including very young children, which greatly expands the potential audience, leading to high revenues.
So, why is the most produced rating R?
Probably because films are also made for artistic reasons, and many directors and crew members are willing to create a film that, while not having the most profitable rating, satisfies their artistic needs.

### vs Budget

Finally, let’s analyze which film ratings receive the most investment.

```{r}
# Calculate the average budget by movie certification
certification_budget <- movie_data |> 
  group_by(certification) |> 
  summarise(avg_budget = mean(budget, na.rm = TRUE)) |> 
  mutate(certification = factor(certification,
                                levels = c("1", "2", "3", "4", "5", "6"),
                                labels = c("G", "PG", "PG-13", "R", "NC-17", "NR")))

# Create a bar chart showing the average budget by certification
ggplot(certification_budget, aes(x = certification, y = avg_budget)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Average Budget by Certification",
       x = "Certification",
       y = "Average Budget ($)") +
  theme_minimal() +
  scale_y_continuous(labels = function(x) paste0(x / 1e6, "M")) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )

# Remove the certification_budget variable after use
rm(certification_budget)
```

It is observed that films with the highest investments are those with a relatively low rating.
These films, compared to those with a higher rating, allow for reaching a wider potential audience.
Consequently, film production companies and investors are more inclined to invest their money in these projects.

## Genres

### Genre Count

To begin, let’s count the number of films for each genre.
It’s important to remember that a film can belong to more than one genre.
Therefore, we expect to find a higher number of films in more general genres, such as *Drama* or *Action*, which can include more specific genres like *Horror*.

```{r}
# Count of movies by genre
# Select the genre columns and transform the data into a long format
genre_counts <- movie_data |> 
  select(Comedy, Action, Crime, Thriller, Adventure, Science_Fiction, 
         Animation, Family, Drama, Romance, Mystery, Horror, Fantasy, 
         War, Music, Western, History, Documentary, TV_Movie) |> 
  pivot_longer(cols = everything(), names_to = "Genre", values_to = "Count") |> 
  mutate(Genre = str_replace_all(Genre, "_", " ")) |>  # Replace underscores with spaces in genre names
  group_by(Genre) %>%
  summarise(Total = sum(Count)) %>%  # Sum the counts for each genre
  arrange(desc(Total))  # Sort genres by total count in descending order

# Plot a bar chart of the number of movies per genre
ggplot(genre_counts, aes(x = reorder(Genre, -Total), y = Total)) +
  geom_bar(stat = "identity", fill = "chocolate") +
  theme_minimal() +
  labs(title = "Number of movies per genre",
       x = "Genre",
       y = "Number of movies") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  ) +
  coord_flip()  # Flip the coordinates for better readability

# Remove the temporary genre_counts data
rm(genre_counts)
```

Indeed, we notice that the three most common genres are *Drama*, *Comedy*, and *Action*.
As mentioned earlier, being very broad genres, they can also encompass more specific genres.

### vs Revenue

Let’s now examine the relationship between different film genres and the revenue generated.

```{r}
# Transform the data into long format for genre analysis
movie_genres_long <- movie_data |> 
  pivot_longer(cols = Comedy:TV_Movie, names_to = "genre", values_to = "is_genre") |> 
  filter(is_genre == 1) |>  # Consider only movies that belong to a specific genre
  mutate(genre = str_replace_all(genre, "_", " "))  # Replace underscores with spaces in genre names

# Box plot of revenue by genre
ggplot(movie_genres_long, aes(x = genre, y = revenue)) +
  geom_boxplot(color = "darkred") +
  scale_y_log10(labels = function(x) paste0(x / 1e6, "M")) +  # Log scale for revenue, labeled in millions
  labs(title = "Distribution of Revenue by Genre",
       x = "Genre", 
       y = "Revenue (log scale)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate genre names on the x-axis
        axis.title = element_text(size = 10))

# Remove the temporary movie_genres_long data
rm(movie_genres_long)
```

The graph clearly shows that genres such as Action, Adventure, Science Fiction, Animation, and Family are generally associated with higher revenues, with a wider distribution of earnings.
On the contrary, as one might expect, genres like Documentary and TV Movies tend to generate lower and less variable revenues.
This suggests that the financial success of a film can be strongly influenced by its genre, with some genres offering greater opportunities for high earnings compared to others.

### vs Revenue/Budget Ratio

Let’s now consider the relationship between film genres and the ratio of revenue to budget.
It is possible that some genres, despite having a lower average revenue in absolute terms, present a higher revenue-to-budget ratio.

```{r}
# List of genres
genres <- c("Comedy", "Action", "Crime", "Thriller", "Adventure", "Science_Fiction", "Animation", "Family", "Drama", "Romance", "Mystery", "Horror", "Fantasy", "War", "Music", "Western", "History", "Documentary", "TV_Movie")

# Calculate the average revenue/budget ratio for each genre
genre_data <- genres %>%
  map_df(function(genre) {
    movie_data %>%
      filter(!!sym(genre) == 1) %>%
      summarise(avg_ratio = mean(rev_bud_ratio, na.rm = TRUE)) %>%
      mutate(genre = str_replace_all(genre, "_", " "))  # Replace underscores with spaces in genre names
  })

# Sort the data by avg_ratio in descending order
genre_data <- genre_data %>% arrange(desc(avg_ratio))

# Create a bar chart of average revenue/budget ratio by genre
ggplot(genre_data, aes(x = avg_ratio, y = reorder(genre, avg_ratio))) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Average Revenue/Budget Ratio by Genre",
       x = "Revenue/Budget Ratio",
       y = "Genre") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        axis.text.x = element_text(hjust = 1),
        axis.title = element_text(size = 10))

# Remove the temporary genre_data and genres variables
rm(genre_data)
rm(genres)
```

In fact, we notice that the genre with the highest ratio is Documentary, followed by more specific genres like Horror and Music.
These films tend to have lower budgets compared to genres like Drama or Action.
This may be due to the fact that the highest-paid stars often prefer to work in more popular genres, such as those mentioned, significantly increasing the film’s budget.

## Cast

Regarding the variable related to the cast, we will specifically analyze the gender of the actors.
In particular, we will conduct an analysis on the possible combinations of the two main actors in the film: male-male, male-female, female-male, and female-female.
For each combination, we will examine the number of films made, the average revenue, the average budget, and the average revenue-to-budget ratio.

### Gender Analysis

```{r}
# Function to convert numerical values into gender initials
gender_to_initial <- function(x) {
  ifelse(x == 1, "M", ifelse(x == 0, "F", as.character(x)))
}

# Prepare data for analyzing star gender combinations
gender_data <- movie_data %>%
  mutate(
    combination = paste(
      gender_to_initial(Star_gender1),  # Convert the first star's gender to initials
      gender_to_initial(Star_gender2),  # Convert the second star's gender to initials
      sep = "-"
    )
  ) %>%
  group_by(combination) %>%
  summarise(
    count = n(),  # Count the number of movies for each combination
    avg_revenue = mean(revenue, na.rm = TRUE),  # Calculate the average revenue
    avg_budget = mean(budget, na.rm = TRUE),  # Calculate the average budget
    avg_rev_bud_ratio = mean(rev_bud_ratio, na.rm = TRUE)  # Calculate the average revenue/budget ratio
  ) %>%
  pivot_longer(cols = c(count, avg_revenue, avg_budget, avg_rev_bud_ratio),
               names_to = "metric", values_to = "value")  # Reshape the data into a long format

# Create a plot with specific colors for each metric and no legend
ggplot(gender_data, aes(x = combination, y = value, fill = metric)) +
  geom_bar(stat = "identity", color = "black") +
  facet_wrap(~ metric, scales = "free_y", nrow = 2, ncol = 2,  # Split the plot by metric, with separate y-scales
             labeller = labeller(metric = c(
               "count" = "Number of movies",
               "avg_revenue" = "Average Revenue",
               "avg_budget" = "Average Budget",
               "avg_rev_bud_ratio" = "Average Rev/Bud Ratio"
             ))) +
  theme_minimal() +
  labs(
    title = "Metrics for Star Gender Combination",
    x = "Gender Combination",
    y = ""
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate gender combinations for readability
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold"),
    legend.position = "none"  # Remove the legend
  ) +
  scale_x_discrete(limits = c("M-M", "M-F", "F-F", "F-M")) +  # Set the order of the gender combinations
  scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +  # Format y-axis labels
  scale_fill_manual(values = c(
    "avg_rev_bud_ratio" = "purple",  # Darker purple for avg_rev_bud_ratio
    "avg_revenue" = "darkred",       # Darker red for avg_revenue
    "avg_budget" = "darkblue",       # Darker blue for avg_budget
    "count" = "darkgreen"            # Darker green for count
  ), guide = "none")  # Disable the legend

# Clean up temporary variables
rm(gender_data)
rm(gender_to_initial)
```

In general, we note that the presence of a male actor among the protagonists is associated with a higher number of films, a higher average revenue, and a higher average budget.
However, the films that present the highest revenue-to-budget ratio are those in which the main star is female and the secondary star is male.

## Crew

### Director gender

Let’s now examine the differences that emerge when a director is male compared to when a director is female.
In particular, we will analyze the number of films made by male and female directors, their average revenue, the average budget, and finally the average revenue-to-budget ratio.

```{r}
# Function to convert numeric values to gender initials
gender_to_initial <- function(x) {
  ifelse(x == 1, "Male", ifelse(x == 0, "Female", as.character(x)))
}

# Data preparation for director analysis
director_data <- movie_data %>%
  mutate(Director_gender = gender_to_initial(Director_gender)) %>%
  group_by(Director_gender) %>%
  summarise(
    count = n(),  # Number of movies for each director gender
    avg_revenue = mean(revenue, na.rm = TRUE),  # Average revenue per gender
    avg_budget = mean(budget, na.rm = TRUE),  # Average budget per gender
    avg_rev_bud_ratio = mean(rev_bud_ratio, na.rm = TRUE)  # Average revenue/budget ratio per gender
  ) %>%
  pivot_longer(cols = c(count, avg_revenue, avg_budget, avg_rev_bud_ratio),
               names_to = "metric", values_to = "value")

# Plotting metrics for director gender
ggplot(director_data, aes(x = Director_gender, y = value, fill = metric)) +
  geom_bar(stat = "identity", color = "black") +
  facet_wrap(~ metric, scales = "free_y", nrow = 2, ncol = 2, 
             labeller = labeller(metric = c(
               "count" = "Number of movies",
               "avg_revenue" = "Average Revenue",
               "avg_budget" = "Average Budget",
               "avg_rev_bud_ratio" = "Average Rev/Bud"
             ))) +
  theme_minimal() +
  labs(
    title = "Metrics for Director Gender",
    x = "Director Gender",
    y = ""
  ) +
  theme(
    axis.text.x = element_text(hjust = 0.5),  # Center-align x-axis labels
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold"),
    legend.position = "none"  # Remove legend
  ) +
  scale_x_discrete(limits = c("Male", "Female")) +
  scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  scale_fill_manual(values = c(
    "avg_rev_bud_ratio" = "purple",  # Darker RdPu color
    "avg_revenue" = "darkred",        # Darker Reds color
    "avg_budget" = "darkblue",         # Darker YlGnBu color
    "count" = "darkgreen"              # Darker YlGn color
  ))

# Cleaning up temporary data
rm(gender_to_initial)
rm(director_data)
```

We can note that the number of films made by female directors is significantly lower compared to those directed by male directors.
Additionally, although less evident, female directors show lower values for all other metrics.

Let’s now see which film genre is most produced by male directors and by female directors.

```{r}
# Function to convert numeric gender codes to initials
gender_to_initial <- function(gender) {
  ifelse(gender == 2, "M", ifelse(gender == 1, "F", NA))
}

# List of movie genres
movie_genres <- c("Comedy", "Action", "Crime", "Thriller", "Adventure", 
                  "Science_Fiction", "Animation", "Family", "Drama", 
                  "Romance", "Mystery", "Fantasy", "War", "Music", 
                  "Western", "History", "Horror", "Documentary", "TV_Movie")

# Determine the most common genre for each director gender
most_common_genre <- movie_data %>%
  mutate(Director_gender = gender_to_initial(Director_gender)) %>%
  select(Director_gender, all_of(movie_genres)) %>%
  pivot_longer(cols = all_of(movie_genres), names_to = "movie_genre", values_to = "is_genre") %>%
  filter(is_genre == 1) %>%
  group_by(Director_gender, movie_genre) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(Director_gender) %>%
  slice_max(order_by = count, n = 1)

# Display the results in a readable format
for(gender in c("M", "F")) {
  genre <- most_common_genre %>% 
    filter(Director_gender == gender) %>% 
    pull(movie_genre)
  
  gender_label <- if(gender == "M") "male directors" else "female directors"
  
  cat(sprintf("The most common film genre for %s: %s\n", gender_label, genre))
}

rm(gender)
rm(genre)
rm(gender_label)
rm(most_common_genre)
```

Below is the code to determine, based on the director’s gender, which film genre generates the highest revenue.

```{r}
# Calculate the genre with the highest revenue for each director gender
highest_revenue <- movie_data %>%
  mutate(Director_gender = gender_to_initial(Director_gender)) %>%
  select(Director_gender, revenue, all_of(movie_genres)) %>%
  pivot_longer(cols = all_of(movie_genres), names_to = "movie_genre", values_to = "is_genre") %>%
  filter(is_genre == 1) %>%
  group_by(Director_gender, movie_genre) %>%
  summarise(avg_revenue = mean(revenue, na.rm = TRUE), .groups = "drop") %>%
  group_by(Director_gender) %>%
  slice_max(order_by = avg_revenue, n = 1)

# Display the results for the highest revenue genre
for(gender in c("M", "F")) {
  genre <- highest_revenue %>% 
    filter(Director_gender == gender) %>% 
    pull(movie_genre)
  
  gender_label <- if(gender == "M") "male directors" else "female directors"
  
  cat(sprintf("The film genre with the highest revenue for %s: %s\n", gender_label, genre))
}

rm(gender_to_initial)
rm(highest_revenue)
rm(movie_genres)
rm(gender)
rm(gender_label)
rm(genre)
```

### Screenplay gender

Let’s now examine the situation related to screenwriting, analyzing the differences between various genres.

```{r}
# Function to convert numeric values to gender initials for screenwriter analysis
gender_to_initial <- function(x) {
  ifelse(x == 2, "Male", ifelse(x == 1, "Female", as.character(x)))
}

# Data preparation for screenwriter analysis
screenplay_data <- movie_data %>%
  mutate(Screenplay_gender = gender_to_initial(Screenplay_gender)) %>%
  group_by(Screenplay_gender) %>%
  summarise(
    count = n(),  # Number of movies for each screenwriter gender
    avg_revenue = mean(revenue, na.rm = TRUE),  # Average revenue per gender
    avg_budget = mean(budget, na.rm = TRUE),  # Average budget per gender
    avg_rev_bud_ratio = mean(rev_bud_ratio, na.rm = TRUE)  # Average revenue/budget ratio per gender
  ) %>%
  pivot_longer(cols = c(count, avg_revenue, avg_budget, avg_rev_bud_ratio),
               names_to = "metric", values_to = "value")

# Custom function to format numbers with units
format_number <- function(x) {
  case_when(
    x >= 1e9 ~ paste0(round(x / 1e9, 1), "B"),  # Format billions
    x >= 1e6 ~ paste0(round(x / 1e6, 1), "M"),  # Format millions
    x >= 1e3 ~ paste0(round(x / 1e3, 1), "K"),  # Format thousands
    TRUE ~ as.character(round(x, 1))
  )
}

# Plotting metrics for screenwriter gender
ggplot(screenplay_data, aes(x = Screenplay_gender, y = value, fill = metric)) +
  geom_bar(stat = "identity", color = "black") +
  facet_wrap(~ metric, scales = "free_y", nrow = 2, ncol = 2, 
             labeller = labeller(metric = c(
               "count" = "Number of movies",
               "avg_revenue" = "Average Revenue",
               "avg_budget" = "Average Budget",
               "avg_rev_bud_ratio" = "Average Rev/Bud"
             ))) +
  theme_minimal() +
  labs(
    title = "Metrics for Screenwriter Gender",
    x = "Screenwriter Gender",
    y = ""
  ) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),  # Center-align x-axis labels
    panel.grid.major.x = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
    panel.grid.minor.x = element_blank(),
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold"),
    legend.position = "none"  # Remove legend
  ) +
  scale_x_discrete(limits = c("Male", "Female")) +
  scale_y_continuous(labels = format_number) +
  scale_fill_manual(values = c(
    "avg_rev_bud_ratio" = "purple",  
    "avg_revenue" = "darkred",        
    "avg_budget" = "darkblue",         
    "count" = "darkgreen"               
  ))

# Cleaning up temporary data
rm(screenplay_data)
rm(format_number)
rm(gender_to_initial)
```

Even in this case, we notice that films written by women are significantly fewer in number compared to those written by men.
However, a particular trend emerges: the average revenue of films written by women is higher than that of films written by men.

## Original Music

### Mean

Let’s move on to examine the percentage of films in our dataset that have an original composer, and therefore a main soundtrack, compared to those that use non-original music and songs.

```{r}
# Calculate the percentage of movies with original music
mean_om = mean(movie_data$original_music, na.rm = TRUE)
cat("The ", round(mean_om, 2) * 100, "% of movies has original music", sep = "")

rm(mean_om)
```

We analyze the average revenue of films with original music compared to those without original music.

```{r}
# Calculate average revenue based on whether movies have original music
revenue_by_music <- movie_data |> 
  group_by(original_music) |> 
  summarise(mean_revenue = mean(revenue, na.rm = TRUE)) |> 
  mutate(music_status = ifelse(original_music == 1, 
                                    "Original Music", 
                                    "Not Original Music"))

# Plotting average revenue for movies with or without original music
ggplot(revenue_by_music, aes(x = music_status, y = mean_revenue, fill = music_status)) +
  geom_bar(stat = "identity", width = 0.7) +
  labs(x = "Music Status", 
       y = "Average Revenue", 
       title = "Average revenue for movies with original music") +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M")) +
  scale_fill_manual(values = c("Not Original Music" = "skyblue", 
                               "Original Music" = "darkblue")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        axis.text.x = element_text(hjust = 0.5),
        axis.title = element_text(size = 10))

rm(revenue_by_music)
```

We notice that films with an original composer, on average, earn much more than those that do not have one.

## CPI

### vs Revenue and Year

Finally, we have included an economic variable in our dataset, the CPI (Consumer Price Index), assuming that the economic situation of a country can influence the number of people going to the cinema, since cinema is not a necessary good.
For this purpose, we graphically compare the CPI index with the revenue, to see if the trend of one over time is similar to the trend of the other.

```{r}
# Calculate the average revenue per year
revenue_by_year <- movie_data %>%
  group_by(year) %>%
  summarise(mean_revenue = mean(revenue, na.rm = TRUE))

# Calculate the average CPI per year (assuming there is only one CPI value per year)
cpi_by_year <- movie_data %>%
  group_by(year) %>%
  summarise(mean_cpi = mean(average_cpi, na.rm = TRUE))

# Combine the two datasets
combined_data <- revenue_by_year %>%
  full_join(cpi_by_year, by = "year") %>%
  pivot_longer(cols = c(mean_revenue, mean_cpi),
               names_to = "metric",
               values_to = "value")

# Create facet plots for the metrics
ggplot(combined_data, aes(x = year, y = value)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ metric, ncol = 1, scales = "free_y", 
             labeller = labeller(metric = c("mean_revenue" = "Average Revenue", 
                                            "mean_cpi" = "CPI Index"))) +
  labs(x = "Year", y = "") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(combined_data$year), 
                                  max(combined_data$year), 
                                  by = 10)) +
  scale_y_continuous(labels = function(x) {
    ifelse(x >= 1e6, paste0(x/1e6, "M"), x)
  }) +
  theme(strip.text = element_text(size = 12, face = "bold"))

# Remove intermediate datasets from memory
rm(combined_data)
rm(cpi_by_year)
rm(revenue_by_year)
```

We can observe that there is indeed a certain similarity in the trends of the CPI and the revenue.

## Correlation Matrix

Let’s conclude this descriptive analysis by showing the correlation matrix between the variables.
We opted for a matrix that shows only those variables that are correlated with other variables with a correlation index of at least 0.30 in absolute value.

```{r}
# Select only the numeric variables from the dataset
numeric_data <- movie_data |>  
  select(where(is.numeric))

# Compute the correlation matrix
cor_matrix <- cor(numeric_data, use = "complete.obs")

# Identify rows with at least one correlation > 0.30 or < -0.30
rows_to_keep <- apply(cor_matrix, 1, function(x) any(abs(x) > 0.30 & abs(x) < 1))

# Filter the matrix to keep only these rows and columns
cor_matrix_filtered <- cor_matrix[rows_to_keep, rows_to_keep]

# Set correlations between -0.30 and 0.30 to NA for visualization purposes
cor_matrix_filtered[abs(cor_matrix_filtered) <= 0.30] <- NA

# Visualize the filtered correlation matrix
corrplot(cor_matrix_filtered, method = "color", na.label = " ", tl.cex = 0.4, number.cex = 0.3, 
         addCoef.col = "black", diag = FALSE, tl.col = "black")

# Remove intermediate datasets from memory
rm(numeric_data)
rm(cor_matrix)
rm(rows_to_keep)
rm(cor_matrix_filtered)
```

Observing the matrix, a strong positive correlation is noted between a movie's budget and its revenue, suggesting that films with higher budgets tend to earn greater box office returns.
This link is particularly evident, as indicated by the dark color of the corresponding cell.
Another significant correlation is between popularity and the average rating received by movies in which the actors have participated, which makes sense since more popular films typically result in higher ratings.
On the other hand, some negative correlations are observed, such as between the "Family" genre and the certification variable, suggesting that films of this genre indeed have a lower rating.
Moreover, the correlation between variables like the number of words in the title and indicators of popularity or financial success appears to be relatively weak or non-existent, highlighting that the length of the title is not a determining factor for a film's success.

# Statistical Test

Regarding statistical tests, we will adopt a different approach.
In this case, we will not proceed from the first to the last variable of the dataset, but we will divide the analysis based on the type of test performed.

## T-test

### IP

```{r}
# Perform the t-test for revenue based on IP (Intellectual Property)
t_test_result <- t.test(revenue ~ IP, data = movie_data)

# Display the t-test results
t_test_result
```

The Welch's t-test performed on the revenue variable in relation to the Intellectual Property (IP) variable compares the average revenues between two groups: films not based on pre-existing intellectual properties (group 0) and those that are (group 1).
The result shows a t-value of -5,2728., indicating a significant difference between the two groups.
The associated p-value is extremely low (1.363e-07), well below the common significance threshold of 0.05.
This suggests that the difference in average revenues between IP-based films and non-IP-based films is highly significant.

The estimates of average revenues show that non-IP based films (group 0) have an average revenue of about \$61 million, while IP-based films (group 1) have a higher average revenue, approximately \$80 million.
The difference between these two groups is further highlighted by the 95% confidence interval, which ranges from approximately -\$26 million to -\$12 million.
Since the confidence interval does not include zero, it confirms that there is a significant difference in average revenues between the two groups.

The image below shows a box plot comparing the revenue distribution between original films and films based on existing intellectual property (IP).

```{r}
# Plot the revenue comparison between original and IP-based movies
ggplot(movie_data, aes(x = factor(IP), y = revenue)) +
  geom_boxplot() +
  scale_y_log10(labels = scales::comma) +
  labs(
    x = "IP", 
    y = "Revenue (log scale)",
    title = "Revenue Comparison Between Original and IP-Based Movies"
  ) +
  scale_x_discrete(labels = c("0" = "Original", "1" = "IP-Based")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )
```

The provided boxplot shows the distribution of movie revenues divided into two groups: Original, representing films not based on pre-existing intellectual properties, and IP-Based, representing films based on existing IPs such as franchises, sequels, or adaptations of well-known works.
The y-axis is presented in a logarithmic scale, which helps to better visualize the distribution of the data, especially when the revenues vary over a wide range.
From the visualization of the boxplot, it is evident that IP-Based films tend to have higher average revenues compared to Original films.
This is indicated by the position of the median line (which represents the median revenue value) being higher in the boxplot for IP-Based films.
Additionally, the box itself, which contains 50% of the data (between the first and third quartiles), is slightly higher for IP-Based films, suggesting that the central distribution of revenues for these films is also higher.
Another interesting aspect of the chart is the presence of numerous outliers (points outside the "whiskers" of the boxplot), especially for the Original films.
These outliers indicate that there are original films that have generated particularly low revenues, much lower than most of the films in the dataset.

### Original Music

```{r}
# Perform the t-test for revenue based on original music presence
t_test_result <- t.test(revenue ~ original_music, data = movie_data)

# Calculate the effect size (Cohen's d)
cohens_d <- effsize::cohen.d(revenue ~ original_music, data = movie_data)

# Visualization of revenue distribution based on the presence of original music
ggplot(movie_data, aes(x = factor(original_music), y = revenue)) +
  geom_boxplot() +
  scale_y_log10(labels = scales::comma) +
  labs(title = "Revenue Distribution for Presence of Original Music",
       x = "Original Music (0 = No; 1 = Yes)", 
       y = "Revenue (log scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )
```

The boxplot displayed highlights the distribution of movie revenues based on the presence or absence of original music.
The original_music variable is coded as 0 for films that do not have original music and 1 for those that do.
The y-axis, as in the previous case, is represented in a logarithmic scale to handle the wide range of revenues present in the data.
From the visual analysis of the boxplot, we notice that films with original music (indicated by 1) tend to have slightly higher revenues compared to those without original music (indicated by 0).
However, the difference does not seem to be particularly pronounced: the medians of both groups are close, indicating that the typical revenue is similar between the two groups.
The boxes, which represent the interquartile range (central 50% of the data), have a similar height, suggesting that the variability of revenues within each group is comparable.
There are, however, some more noticeable outliers in the group without original music, indicating that there are films without original music that have had particularly low revenues, much lower compared to most other films.
Despite this, the overall distribution seems to indicate that the presence of original music might have a positive effect on revenues, even if it is not a marked difference.

```{r}
# Generate descriptive statistics
summary_stats <- movie_data |> 
  group_by(original_music) |> 
  summarise(
    mean_revenue = mean(revenue, na.rm = TRUE),
    median_revenue = median(revenue, na.rm = TRUE),
    sd_revenue = sd(revenue, na.rm = TRUE),
    n = n()
  )

# Print the t-test results
print("T-test Result:")
t_test_result

# Print the effect size (Cohen's d)
print("Effect Size (Cohen's d):")
cohens_d

# Remove intermediate datasets from memory
rm(summary_stats)
rm(t_test_result)
rm(cohens_d)
```

The results of the Welch's t-test and the effect size measure (Cohen's d) indicate that there is a statistically significant difference in average revenues between films with and without original music.
The t-test produced an extremely high t-value (-24.684) and a p-value less than 2.2e-16, confirming that the difference in average revenues between the two groups is not due to chance.
Films with original music (group 1) have a significantly higher average revenue, approximately \$78.4 million, compared to films without original music (group 0), which have an average revenue of about \$19.4 million.
The 95% confidence interval for the difference between the means does not include zero, further strengthening the significance of the difference.
However, a Cohen's d of -0.40 indicates that, although the difference is statistically significant, the magnitude of the effect is small.
This suggests that while the presence of original music is associated with higher revenues, the effect of this variable on revenues is not very large in terms of standard deviation units.
Thus, original music can contribute to a film's financial success, but it might not be a principal determining factor.

## ANOVA

### Certification

```{r}
# ANOVA for revenue based on certification
revenue_anova <- movie_data %>% 
  mutate(certification = as.factor(certification)) %>% 
  {aov(revenue ~ certification, data = .)}

# Summary of the ANOVA results for revenue
summary(revenue_anova)
```

The analysis of variance (ANOVA) conducted on the dataset assessed whether there are significant differences in the average revenues of films based on the film's certification (a categorical variable with various classifications, such as "G", "PG", "PG-13", "R").
The ANOVA output shows that the certification variable has a significant effect on revenues, as indicated by the high F-value (138.8) and the extremely low p-value (less than 2e-16).
This p-value suggests that the probability that the observed differences in revenues across different certifications are due to chance is virtually nil.
The Sum Sq value for certification reflects the amount of variability in revenues that can be explained by certification, while the Mean Sq is the average variance associated with this variable.
The presence of high statistical significance suggests that there are substantial differences in average revenues among at least some of the certifications considered.
This implies that the film certification might be an important factor to consider in the financial performance of movies, potentially influencing audience reach and revenue generation.

```{r}
# Post-hoc analysis (Tukey HSD) for revenue
TukeyHSD(revenue_anova)

# Remove the ANOVA object from memory
rm(revenue_anova)
```

The post-ANOVA analysis you've described involves a Tukey's Honestly Significant Difference (HSD) test or a similar method to compare the means of revenues among different certification groups.
This type of test is commonly used after an ANOVA to conduct pairwise comparisons between groups to identify which specific certification groups significantly differ from each other in terms of revenues.
The output provides a series of comparisons between certifications (numbered from 1 to 6), showing the average revenue difference (diff), the 95% confidence interval for that difference (lwr and upr), and the adjusted p-value (p adj) for each comparison.
Significant differences in average revenues, as indicated by very low p-values in comparisons like 3-1, 4-1, 5-1, and 6-1, suggest that films with these certifications have significantly different revenues compared to group 1.
For instance, the difference between groups 6 and 1 has a confidence interval that does not include zero and an extremely low p-value, indicating a highly significant difference.
Further, comparisons such as 6-4 and 6-3 show significant differences between certifications 6 and 4, and 6 and 3, respectively, suggesting that there are also relevant differences in average revenues between these groups.
This detailed breakdown can help stakeholders understand how certification influences movie earnings and can guide decisions in production and marketing to optimize financial outcomes based on the target audience’s certification preferences.

### Month

```{r}
# ANOVA for revenue based on the month of release
month_revenue_anova <- movie_data %>%
  mutate(month = as.factor(month)) %>%
  aov(revenue ~ month, data = .)

# Summary of the ANOVA results
summary(month_revenue_anova)
```

The analysis of variance (ANOVA) conducted on the revenues based on the release month of the films suggests that there are significant differences in average revenues depending on the month in which a film is released.
The F-value of 23.89 and the p-value less than 2e-16 indicate that the probability that these differences are due to chance is virtually nil.
This result implies that the release month is a relevant factor for a film's revenues.
The Sum Sq value for the month (5.658e+18) represents the variability in revenues that can be explained by the month variable, while the Mean Sq indicates the average variance associated with the release months.
The fact that the p-value is extremely low and the significance code confirm that the differences between the months are highly significant.

Although ANOVA tells us that there are significant differences, it does not tell us which specific months differ from each other.
For this, it would be necessary to examine the results of Tukey’s post-hoc test.

```{r}
# Levene's test for homogeneity of variances
leveneTest(revenue ~ month, data = movie_data)
```

The Levene's test is used to assess the homogeneity of variances, that is, whether the variances of revenues across different film release months are equal.
The output of the Levene's test indicates that the F-value is 23.109 and the associated p-value is less than 2.2e-16, which is extremely low.
This very low p-value suggests that there is a significant difference in the variances of revenues across different months.
In other words, the test rejects the null hypothesis of equal variances, indicating that the variance in revenues is not consistent across the release months.
This is an important result because ANOVA assumes homogeneity of variances, and a violation of this assumption can affect the reliability of the ANOVA results.

```{r}
# Effect size (eta squared) calculation
eta_squared(month_revenue_anova)
```

The eta squared (η²) is a measure of effect size that indicates the proportion of total variability explained by a particular independent variable in the context of an ANOVA.
In this case, eta squared is calculated to assess the effect of the release month on movie revenues.
The output shows that the eta squared for the variable "month" is 0.03, with a 95% confidence interval ranging from 0.02 to 1.00.
This value of 0.03 means that 3% of the total variability in movie revenues can be attributed to the release month.
Although this effect is statistically significant (as shown by the ANOVA), the magnitude of the effect is relatively small.
In other words, while the release month does affect revenues, this variable explains only a small portion of the total variability.
The confidence interval, which ranges from 0.02 to 1.00, suggests some uncertainty in the estimate, but the upper value of the interval (1.00) is artificially limited since eta squared cannot exceed 1.0.
This indicates that, although there is an effect of the release month, its relevance in terms of explaining the variability in revenues is limited.

```{r}
# Post-hoc analysis (Tukey HSD) for the month-based ANOVA
tukey_results <- TukeyHSD(month_revenue_anova)

# Plot the revenue distribution by month of release
ggplot(movie_data, aes(x = factor(month), y = revenue)) +
  geom_boxplot() +
  scale_y_log10(labels = scales::label_number(scale = 1e-6, suffix = "M")) +
  scale_x_discrete(labels = c("01" = "January", "02" = "February", "03" = "March", 
                              "04" = "April", "05" = "May", "06" = "June",
                              "07" = "July", "08" = "August", "09" = "September", 
                              "10" = "October", "11" = "November", "12" = "December")) +
  labs(title = "Revenue Distribution by Month of Release",
       x = "Month", y = "Revenue (log scale)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10))

# Remove the month-based ANOVA and Tukey results from memory
rm(month_revenue_anova)
rm(tukey_results)
```

The boxplot shows the distribution of movie revenues based on the release month, with the y-axis in logarithmic scale to handle the wide range of revenues.
Each box represents the revenue distribution for films released in a particular month, providing information on values such as the median, the interquartile range, and outliers.
From the chart, it is noticeable that, although there are some variations between months, the median revenue remains fairly similar for most months.
However, some months, like July and December, show a tendency towards higher average revenues compared to other months, likely due to the summer season and holiday periods, which are famously profitable times for movie releases.
Outliers are present in all months, indicating that there are films which, regardless of the release month, have achieved significantly lower or higher revenues compared to most of the films released in the same period.
The presence of these outliers suggests that, apart from the release month, there are other relevant factors influencing movie revenues, such as budget, promotion, cast, and competition.

## Wilcoxon-Mann-Whitney Test

### Cast Gender

```{r}
# Filter the data to remove any NAs and select relevant columns
data_filtered <- movie_data %>%
  filter(!is.na(rev_bud_ratio) & !is.na(Star_gender1) & !is.na(Star_gender2)) %>%
  select(rev_bud_ratio, Star_gender1, Star_gender2)

# Create a new variable to distinguish between movies with male and female stars
data_filtered <- data_filtered %>%
  mutate(gender_category = case_when(
    Star_gender1 == 1 & Star_gender2 == 1 ~ "Both Male",
    Star_gender1 == 0 & Star_gender2 == 0 ~ "Both Female",
    TRUE ~ "Mixed"
  ))

# Perform the Wilcoxon-Mann-Whitney test between movies with "Both Male" and "Both Female" stars
wilcox_test <- wilcox.test(rev_bud_ratio ~ gender_category, 
                           data = data_filtered %>% 
                           filter(gender_category %in% c("Both Male", "Both Female")))

# Display the Wilcoxon test results
wilcox_test

# Remove the filtered data and test results from memory
rm(data_filtered)
rm(wilcox_test)
```

The Wilcoxon-Mann-Whitney test, conducted on the revenue-to-budget ratio (rev_bud_ratio) in relation to the gender of the two main stars of the film, aims to determine if there is a significant difference in the distribution of this ratio between films with both male protagonists ("Both Male") and those with both female protagonists ("Both Female").
The output of the test shows a W value of 1,061,904 and a p-value of 0.4596.
Since the p-value is above the common significance threshold, there is not enough statistical evidence to reject the null hypothesis.
The null hypothesis in this context is that there is no significant difference in the distribution of the revenue-to-budget ratio between the two groups of films.
In other words, the results of the test indicate that there is no statistically significant difference in the revenue-to-budget ratio between films with male protagonists and those with female protagonists.
This suggests that, at least based on this sample and this analysis, the gender of the two main stars does not significantly influence the economic effectiveness of a film in terms of its revenue-to-budget ratio.

# Final Dataset

Since we have to many *screenwriters* without a gender, we are going to remove the columns related to this role.

```{r}
movie_data <- movie_data |> 
  select(-c(Screenplay_rev, 
            Screenplay_gender,
            Screenplay_pop,
            Screenplay_vote))
```

Finally, we remove the `rev_bud_ratio` variable from our dataset.

```{r}
movie_data <- movie_data |> 
  select(-rev_bud_ratio)
```

At this point, we can proceed to save the new dataset, which we will use for future analyses.

```{r}
# Save the final dataset to an Excel file
write.xlsx(movie_data, "final_dataset.xlsx", rowNames = FALSE)
```
